Review of Edge Detectors

### 2012-07-23 edge-detection-review-v8.tex
new file:   1dmasks_n.pdf 	new file:   2dconvolution.tex 	new file:   bibliography.bib 	new file:   blockdiagram1.pdf 	new file:   blockdiagram2.pdf 	new file:   edge-detection-review-v10.tex 	new file:   edge-detection-review-v2.tex 	new file:   edge-detection-review-v3.tex 	new file:   edge-detection-review-v4.tex 	new file:   edge-detection-review-v5.tex 	new file:   edge-detection-review-v6.tex 	new file:   edge-detection-review-v7.tex 	new file:   edge-detection-review-v8.tex 	new file:   edge-detection-review-v9.tex 	new file:   edge-detection-review.tex 	new file:   examples/lena.png 	n

%-------------------------------------------------------------------------------
% A Review of Classic Edge Detectors - v10
% by Haldo Spontón & Juan Cardelino
% IPOL 2012
%-------------------------------------------------------------------------------

@@

\usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{sidecap,array}
\usepackage[hang,center]{subfigure}
\usepackage[table]{xcolor}
\usepackage{multicol}
\usepackage{listings}
\usepackage{color}
%\usepackage{algorithm}
\usepackage{algorithmic}
\newtheorem{theorem}{Theorem}
\numberwithin{equation}{section}
\numberwithin{table}{section}
%\numberwithin{figure}{section}
\begin{document}
\bibliographystyle{ieeetr}
\lstset{
	backgroundcolor=\color[rgb]{0.94,0.94,0.94},
    tabsize=4,
    inputencoding=utf8x,
         extendedchars=\true,
         language=C,
         basicstyle=\scriptsize,
                 showtabs=false,
        showspaces=false,
        showstringspaces=false,
        identifierstyle=\ttfamily,
        keywordstyle=\color[rgb]{0,0,1},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
}

@@

\begin{abstract}
In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
behind edge detection is to find where abrupt changes in the intensity of an image have occurred. 
The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
such as Sobel, Prewitt and Roberts. In the second reviewed family it is used second derivative, for example in algorithms 
like Marr-Hildreth and Haralick. \\
Results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time) are compared, considering different ways to convolve an 
image with a kernel (step required in some of the algorithms).
\end{abstract}

@@

\begin{ipolCode}
For all the algorithms reviewed, an open source C implementation is provided and can be downloaded from 
\href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}. You can also follow 
this code project in this \href{git://github.com/haldos/edges.git}{GIT repository}. An 
\href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{online demonstration} is also available, where 
you can test and reproduce our results.
\end{ipolCode}

@@

In the 70's, edge detection methods were based on using small operators 
(such as Sobel masks), attempting to compute an approximation of the
first derivative of the image \cite{Gonzalez2007Digital}. In section ~\ref{sec:first} the 
behavior of such algorithms is described, as an introduction to a more sophisticated 
analysis of the edge detection process.\\

In 1980 Marr and Hildreth \cite{AIM-518} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or trough) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
which is also reviewed in section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to approximate locally the image around a point. Then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is achieved by imposing a condition over the model parameters. \\

@@

To compute the gradient, the computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
at every pixel in the image is required. When dealing with digital images, numerical approximations 
of the partial derivatives are required, calculated in a neighborhood of each point. In the following,
the methods of Roberts, Prewitt and Sobel will be studied, whose main difference is how they perform this calculation.

@@

Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
\begin{equation*}
	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{th} \\
									1,& \mbox{if}\ M[i]<\mbox{th}
					  \end{cases}
\end{equation*}
where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
It is noted that different thresholds lead to different results. Thicker or more edges for smaller thresholds, fewer edges 
for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
of the gradient image. The advantage of this is to adapt the threshold to the dynamic range of the image. 
Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.

@@

The Marr-Hildreth edge-detection algorithm may be summarized as follows:
\begin{enumerate}
	\item Filter the input image with a $n \times n$ Gaussian lowpass filter obtained by sampling equation \ref{eq:gaussian_function}.
	\item Compute the Laplacian of the image resulting from step 1, using, for example, the $3\times3$ mask:
			$\begin{bmatrix}
			1 &  1 & 1 \\
			1 & -8 & 1 \\
			1 &  1 & 1 \\
			\end{bmatrix}$\footnote{Steps 1 and 2 can be merged into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}
	\item Find the zero crossings of the image from step 2.
\end{enumerate}

@@

The zero-crossing threshold ($th_{ZC}$) is given as a percentage of the maximum value $max_L$ of the Laplacian 
image (both Gaussian and LoG kernels). Each pixel $p$ has eight neighbors, named according to their position 
as follows.

@@

Here, the interpolation method used in the original work of Haralick is presented, although there are other options to do this \cite{getreuer}. \\

@@

To solve this problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest 
neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.
In Haralick's work, this approximation is computed using least squares, and to speed up the calculations the solution is 
computed by convolving the image with some precomputed masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\

@@

The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If just the first order terms of this 
polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approxmated as

@@

where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
should be zero at $\rho=0$, but this condition is relaxed, allowing the the polynomial becomes zero in a neighborhood 
of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
value is between $0.4$ and $0.6$.\\

@@

Given the direction defined by $\theta$, with $\rho>0$, one edge will always be an ascending step. This 
indicates that first derivative in this direction has a maximum and that second derivative has a zero 
crossing with negative slope. That is, third derivative less than zero.

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF
\STATE Define the ten $5x5$ masks ($mask[1]\dots mask[10]$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

The first step in every algorithm presented above is to convert color images to gray intensity images. 
For this purpose, \href{http://www.libpng.org/}{libpng} coefficients are used,
\begin{equation}
    Y = (6968 R + 23434 G + 2366 B) / 32768
\end{equation}
where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float} to avoid truncated results. See code in section \ref{app:marr-hildreth}.}.

@@

Gaussian kernel is generated by sampling the 2-D Gaussian function
%\footnote{Note that for simplicity the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted.}
\begin{equation}
	\label{eq:gaussian_function}
	G(x,y) = \frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{x^2+y^2}{2\sigma^2}}
\end{equation}
where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).\\

The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
input parameters, but these are not strictly independent of each other. 
A value of $n$ large enough is needed to ensure that no information is lost when creating the kernel $G$. To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$. Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.\\

@@

\begin{equation} 
	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}

where the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted. Therefore we obtain

\begin{equation}
	\label{eq:log_function}
	LoG(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
\end{equation}\\

@@

When making a convolution, it is necessary to define the boundary conditions used to calculate such convolution 
around the edges of the image, and to get a valid output the same size as input image. In this paper, 
two methods were implemented: zero-padding and boundary reflection. Zero-padding implies completing 
the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
relative to the edge of the image. Direct convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution 
or recursive filtering. Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.

@@

To the best of our knowledge, the only freely available codes implementing this algorithms are the Matlab implementation of edge detection algorithms using 2D masks (\href{http://www.mathworks.com/help/toolbox/images/ref/edge.html}{\texttt{edge} command}) and a Haralick algorithm implementation in Matlab published in \href{http://www.mathworks.com/matlabcentral/fileexchange/35329-simple-edge-detection-using-classical-haralick-method}{Matlab Central}. There is also an implementation similar to the Marr-Hildreth algorithm, as well available in \href{http://www.mathworks.com/matlabcentral/fileexchange/11572-sdgd-edge-detection-filter}{Matlab Central}. All these links were last checked in July 2012. \\

This code documentation was generated using a perl script to extract some parts of the 
code and comments into latex files. This code documentation tool is available as a \href{https://github.com/juan-cardelino/source_comment_extractor}{GIT project}. \\

Also a Doxygen generated documentation of the functions used is available \href{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1}{here}. \\

Required external libraries: 
\begin{itemize}
	\item \href{http://dev.ipol.im/git/coco/iio.git}{\texttt{iio}}.
	\item \href{http://www.libpng.org/}{\texttt{libpng}}. 
\end{itemize}

@@

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image. The first three (\subref{fig:result1-a}, \subref{fig:result1-b} and \subref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\subref{fig:result1-d} and \subref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\subref{fig:result1-f}) corresponds to the Haralick algorithm. Note the rounded corners on the Marr-Hildreth results, due to Gaussian blur.}
	\label{fig:result1}
\end{figure}

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.040 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including I/O) using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
	\label{exectime1}
\end{table}
\vspace{0.5cm}

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. All algorithms show consistent results. Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm. \\

@@

	\subfigure[Sobel. $th=0.1$.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
	\quad
	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3$, $n=25$, $th_{ZC}=0.07$.]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}

	\subfigure[Marr-Hildreth (LoG). $\sigma=3$, $n=29$, $th_{ZC}=0.13$.]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
	\caption{Results of the first derivative, Marr-Hildreth and Haralick's algorithms using Figure \ref{fig:original2} as input image.}
	\label{fig:result2}
\end{figure}

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including I/O) using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}

@@

The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation. Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels). The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity of the edges. Gaussian filtering would reduce noise, alleviate (but not solve) the discontinuity of the edges, and would make more fair the comparison with the second derivative methods. \\

@@

Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas. It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinomial function. This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too. \\

All algorithms have their advantages and disadvantages. Choosing one or the other may depend on requirements of the application. Furthermore, none of the implemented methods ensure edge connectivity, as the Canny \cite{Canny1986} algorithm does.\\

>>>>

%-------------------------------------------------------------------------------
% A Review of Classic Edge Detectors - v8
% by Haldo Spontón & Juan Cardelino
% IPOL 2012
%-------------------------------------------------------------------------------

@@

\usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{sidecap,array}
\usepackage[hang,center]{subfigure}
\usepackage[table]{xcolor}
\usepackage{multicol}
\usepackage{listings}
\usepackage{color}
%\usepackage{algorithm}
\usepackage{algorithmic}
\newtheorem{theorem}{Theorem}
\numberwithin{equation}{section}
\numberwithin{table}{section}
\numberwithin{figure}{section}
\begin{document}
\bibliographystyle{ieeetr}
\lstset{
	backgroundcolor=\color[rgb]{0.94,0.94,0.94},
    tabsize=4,
    inputencoding=utf8x,
         extendedchars=\true,
         language=C,
         basicstyle=\scriptsize,
                 showtabs=false,
        showspaces=false,
        showstringspaces=false,
        identifierstyle=\ttfamily,
        keywordstyle=\color[rgb]{0,0,1},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
}

@@

\begin{abstract}
In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
behind edge detection is to find where abrupt changes in the intensity of an image have occurred. 
The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
such as Sobel, Prewitt and Roberts. In the second family reviewed, second derivative is used, for example in algorithms 
like Marr-Hildreth and Haralick. \\
Results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time) are compared, considering different ways to convolve an 
image with a kernel (step required in some of the algorithms).
\end{abstract}

@@

\begin{ipolCode}
For all the algorithsm reviewed, an open source C implementation is provided and can be downloaded from 
\href{http://iie.fing.edu.uy/~haldos/downloads/edge_detectors_v0.1.tar.gz}{here}. You can also follow 
this code project in this \href{git://github.com/haldos/edges.git}{GIT repository}. An 
\href{http://dev.ipol.im/~haldos/ipol_demo/xxx_edges/}{online demonstration} is also available, where 
you can test and reproduce our results.
\end{ipolCode}

@@

In the 70's, edge detection methods were based on using small operators 
(such as Sobel masks), attempting to compute an approximation of the
first derivative of the image \cite{Gonzalez2007Digital}. In section ~\ref{sec:first} the 
behavior of such algorithms is described, as an introduction to a more sophisticated 
analysis into the edge detection process.\\

In 1980 Marr and Hildreth \cite{AIM-518} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or trough) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
which is also reviewed in section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to locally approximate the image around a point. Then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is done by imposing a condition over the model parameters. \\

@@

To compute the gradient, the computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
at every pixel in the image is required. When dealing with digital images, numerical approximations 
of the partial derivatives over a neighborhood about a point are required. In the following,
the methods of Roberts, Prewitt and Sobel will be studied, whose main difference is how they perform this calculation.

@@

Finally, the edges image is obteined by thresholding the gradient magnitude image, i.e.\ for each pixel in the image compute
\begin{equation*}
	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{threshold} \\
									1,& \mbox{if}\ M[i]<\mbox{threshold}.
					  \end{cases}
\end{equation*}
A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
It is noted that different thresholds lead to different results. Thicker edges for smaller thresholds, lack of edges 
for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
of the image to be processed. The advantage of this is to adapt the threshold to the dynamic range of the image. 
Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.

@@

The Marr-Hildreth edge-detection algorithm may be summarized as follows:
\begin{enumerate}
	\item Filter the input image with a $n \times n$ Gaussian lowpass filter obtained by sampling equation \ref{eq:gaussian_function}.
	\item Compute the Laplacian of the image resulting from step 1, using, for example, the $3\times3$ mask:
			$\begin{bmatrix}
			1 &  1 & 1 \\
			1 & -8 & 1 \\
			1 &  1 & 1 \\
			\end{bmatrix}$\footnote{Steps 1 and 2 can be made into one, using a $n\times n$ LoG lowpass filter obtained by sampling equation \ref{eq:log_function}.}
	\item Find the zero crossings of the image from step 2.
\end{enumerate}

@@

Zero-crossing threshold ($th_{ZC}$) is given as a percentage of the maximum value $max_L$ of the Laplacian 
image (both Gaussian and LoG kernels). Each pixel $p$ has eight neighbors, named according to their position 
as follows.

@@

Here, the interpolation method used by Haralick is presented, although there are other options to do this \cite{getreuer}. \\

@@

To solve this problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest 
neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.
For speeding up the computation, it is possible to convolve the input image with some precomputed masks, instead 
of finding the least square solution to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\

@@

The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If just the first order terms of this 
polynomial are taken, the gradient angle $\theta$, defined with negative y-axis, can be approxmated as

@@

where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. \\

@@

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF
\STATE Define the ten $5x5$ masks ($mask[1]\dots mask[10]$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{$|C_2/3C_3|\leq\rho_0$}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

The first step in every algorithm presented above is to convert color images to gray intensity images. 
For this purpose, \href{http://www.libpng.org/}{libpng} coefficients are used,
\begin{equation}
    Y = (6968 R + 23434 G + 2366 B) / 32768
\end{equation}
where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float}. See code in section \ref{app:marr-hildreth}.}.

@@

Gaussian kernel is generated by sampling the 2-D Gaussian function\footnote{Note that for simplicity
the normalizing coefficient $1/\sqrt{2\pi\sigma^2}$ is omitted.}
\begin{equation}
	\label{eq:gaussian_function}
	G(x,y) = e^{-\frac{x^2+y^2}{2\sigma^2}}
\end{equation}
where $\sigma$ is the standard deviation (sometimes $\sigma$ is called the \textit{scale space constant}).\\

The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
input parameters, but these are not strictly independent of each other. 
A value of $n$ big enough is needed to ensure that no information is lost when creating the kernel $G$. To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$. Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.\\

@@

\begin{equation} 
	\frac{\partial}{\partial x}G(x,y)=-\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
\end{equation}

Therefore we obtain

\begin{equation}
	\label{eq:log_function}
	LoG(x,y)=\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
\end{equation}\\

@@

When making a convolution, it is necessary to define the boundary conditions used to calculate such convolution 
around the edges of the image, and to get a valid output the same size as input image. In this paper, 
two methods were implemented: zero-padding and boundary reflection. Zero-padding implies completing 
the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
relative to the edge of the image. Convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution 
or recursive filtering. Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.

@@

This code documentation was generated using a perl script to extract some parts of the 
code and comments into latex files. This code documentation tool is available as a GIT
project: 
\begin{center} 
\url{https://github.com/juan-cardelino/source_comment_extractor}. \\
\end{center}
Also a Doxygen generated documentation of the functions used is available from:
\begin{center}
\url{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1} \\
\end{center}
Required external libraries: 
\begin{center}
\texttt{iio} (\url{http://dev.ipol.im/git/coco/iio.git}) \\
\texttt{libpng} (\url{http://www.libpng.org/})
\end{center}

@@

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image. The first three (\ref{fig:result1-a}, \ref{fig:result1-b} and \ref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\ref{fig:result1-d} and \ref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\ref{fig:result1-f}) corresponds to the Haralick algorithm.}
	\label{fig:result1}
\end{figure}

\begin{table}[t!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.040 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including i/o) using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
	\label{exectime1}
\end{table}
\vspace{0.5cm}

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. All algorithms show consistent results. Thicker edges appear in the results of Haralick, due to smooth edges model that is imposed in this algorithm. \\

@@

	\subfigure[Sobel. $th=0.1$.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
	\quad
	\subfigure[Marr-Hildreth (Gaussian). $\sigma=3.5$, $n=25$, $th_{ZC}=0.05$.]{\label{fig:result2-d}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}

	\subfigure[Marr-Hildreth (LoG). $\sigma=3.5$, $n=29$, $th_{ZC}=0.1$.]{\label{fig:result2-e}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result2-f}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
	\caption{Results of the first derivative, Marr-Hildreth and Haralick's algorithms using Figure \ref{fig:original2} as input image.}
	\label{fig:result2}
\end{figure}

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including i/o) using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}

@@

The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation. Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels). The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity of the edges. \\

@@

Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas. It is the only one of these algorithms which works with an approximation of the intensity in the neighborhood of a pixel, using a bicubic polinimial function. This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too. \\

All algorithms have their advantages and disadvantages. Choosing one or the other may depend on requirements of the application. \\
